---
title: ""
author: ""
date: ""
output: html_document
css: C:/ncstate/715/finalproject/nps-fire-viz/markdown_CSS.css
---

<style type="text/css">
.main-container {
  max-width: 1800px;
  margin-left: auto;
  margin-right: auto;
}
</style>

#Visualizing Fire Regime Trends in Everglades National Park and Big Cypress National Preserve
_Nikki Inglis_  
_GIS715 April 2019_    
\  
\    
\    
\   
\  
\  
\  
\  

     
###**Explore the spatial nature of the Everglades fire regime by time of year** 
#### Use the brush tool by clicking on the color below. On the bubble chart, click and drag the time frame you want to explore.  
```{r setup, include=FALSE}
library(ggplot2)
library(ggthemes)
library(scales)
library(maps)
library(mapproj)
library(sp)
library(rgdal)
library(spatialEco)
library(tidyverse)
library(sf)
library(plotly)
library(Cairo)
library(ggspatial)
library(leaflet)

fires<-readOGR("C:/ncstate/NPS/southflorida/allfires/all_fires_cost.shp")

everWFMI<-read.csv("C:/ncstate/NPS/southflorida/adddates/ever_dates.csv", stringsAsFactors = F)
bicyWFMI<-read.csv("C:/ncstate/NPS/southflorida/adddates/BICY_dates.csv", stringsAsFactors = F)
WFMI<-rbind(everWFMI, bicyWFMI)
fires<-merge(fires, WFMI, by.x = "FireID", by.y = "FireId", all.x= T)

fires$StartDate<-as.character(fires$StartDate)
fires$ContrDate<-as.character(fires$ContrDate)
fires$StartDate<-as.Date(fires$StartDate, format="%Y%m%d", origin="1970-01-01")
fires$ContrDate<-as.Date(fires$ContrDate, format="%Y%m%d", origin="1970-01-01")
fires$duration<-difftime(fires$ContrDate, fires$StartDate, units="days")
fires$plot_date<-as.Date(fires$StartDate, format = "%Y%m%d")
fires$plot_date<-as.Date(format(fires$plot_date,"2017-%m-%d"))
fires$NEWCAT<-case_when(fires$FireCause == 'Natural' ~ "Natural wildfire",
                        fires$FireTypePr == 48 ~ "Prescribed Fire",
                        fires$FireTypePr != 48 & fires$FireCause == 'Human' ~ "Human-caused wildfire")
fires<-subset(fires, GISAcres>2000)
fires<-spTransform(fires, '+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs')
firessf<-st_as_sf(fires)
pal<-c()
```

```{r first, echo=F, warning=F, message=F, fig.width=16, fig.height=16}
 library(crosstalk)

share<-SharedData$new(firessf)

bscols(
  ggplotly(
    ggplot(share) + 
    geom_sf(aes(fill=NEWCAT, group=1, text=paste('Name:', IncidentNa, '<br>Type:', NEWCAT, '<br>Date:', format(StartDate, "%b %d %Y"), '<br>Duration:', duration, "days", "<br>Size:", format(round(as.numeric(GISAcres), 0), big.mark=","), "acres")), alpha=0.5) +
    xlab("") +
    ylab("") +
    theme(plot.background = element_rect(fill='gray20'), 
          panel.background = element_rect(fill='gray20'), 
          axis.text = element_blank(),
          panel.grid.major = element_line(colour = "gray20"),   
          panel.grid.minor = element_blank(),
          legend.position='none') +
    scale_fill_manual(values=c("#FF5733", "#ffff00","#56B4E9")), tooltip="text") %>%
    highlight(dynamic = TRUE, persistent = TRUE, color=c('#ff00ff','#fce897','#e98756', '#33ffc2')),
  ggplotly(ggplot(share, aes(y=FIRE_YEAR)) +
    geom_hline(yintercept = seq(1992, 2015, by = 1), color = "gray", size = 0.04) +
    scale_size_area(max_size = 15, guide = FALSE) +
    scale_x_date(date_breaks = "months", date_labels = "%b") +
    scale_y_reverse(limits = c(2015,1992), breaks = c(2010,2005,2000,1995)) +
    xlab("") +
    ylab("") +
    theme(plot.background = element_rect(fill='gray20'), 
          panel.background = element_rect(fill='gray20'), 
          axis.text.y = element_text(colour='gray98', size=16),
          axis.text.x = element_text(colour='gray98', size=10),
          panel.grid.major = element_blank(),   
          panel.grid.minor = element_blank(),
          panel.grid.major.y = element_line(size=1.2, color="gray"),
          axis.ticks.x = element_line(colour='gray98'),
          legend.text = element_text(colour='gray98'),
          legend.title = element_blank(),
          legend.background = element_rect(fill='gray20')) +
    geom_point(aes(size = GISAcres*2, x = plot_date, color = NEWCAT, group=1, text=paste('Name:', IncidentNa, '<br>Type:', NEWCAT, '<br>Date:', format(StartDate, "%b %d %Y"), '<br>Duration:', duration, "days", "<br>Size:", format(round(as.numeric(GISAcres), 0), big.mark=","), "acres")), alpha=0.5) +
    scale_color_manual(values=c("#FF5733", "#ffff00","#56B4E9")), tooltip="text") %>%
    highlight("plotly_selected", persistent = FALSE, color=c('#ff00ff','#fce897','#e98756', '#33ffc2')),
    widths=c(6,6))
```

&nbsp;

&nbsp;

&nbsp;


###**When are most prescribed fires conducted? What month has the most concentration of wildfires?**
####Hover over a color on either chart below to explore how each fire type is distributed throughout history and across the year. The second chart is a density plot of fire occurences over the calendar year. 
```{r second, echo=F, warning=F, message=F, fig.width=14, fig.height=12, fig.align='center'}
setwd("C:/ncstate/NPS/southflorida/")

library(dplyr)
library(readr)
library(ggplot2)
library(ggthemes)
library(scales)
library(maps)
library(mapproj)
library(sp)
library(rgdal)
library(ggmap)

everWFMI<-read.csv("adddates/ever_dates.csv", stringsAsFactors = F)
bicyWFMI<-read.csv("adddates/BICY_dates.csv", stringsAsFactors = F)
allfires<-read.csv("allfires/allfires_vizproject_nodates.csv", stringsAsFactors = F)
WFMI<-rbind(everWFMI, bicyWFMI)
fires_ns<-merge(allfires, WFMI, by.x = "FireID", by.y = "FireId", all.x= T)
cost<-read.csv("2010-17/2010-2017_code_ID_combo.csv")
fires_ns<-merge(fires_ns, cost, by.x = "FireID", by.y = "fireID", all.x=T)
fires_ns$StartDate<-as.character(fires_ns$StartDate)
fires_ns$ContrDate<-as.character(fires_ns$ContrDate)
fires_ns$StartDate<-as.Date(fires_ns$StartDate, format="%Y%m%d", origin="1970-01-01")
fires_ns$ContrDate<-as.Date(fires_ns$ContrDate, format="%Y%m%d", origin="1970-01-01")
fires_ns$duration<-difftime(fires_ns$ContrDate, fires_ns$StartDate, units="days")
fires_ns$plot_date<-as.Date(fires_ns$StartDate, format = "%Y%m%d")
fires_ns$plot_date<-as.Date(format(fires_ns$plot_date,"2017-%m-%d"))
fires_ns$NEWCAT<-case_when(fires_ns$FireCause == 'Natural' ~ "Natural wildfire",
                        fires_ns$FireTypePr == 48 ~ "Prescribed Fire",
                        fires_ns$FireTypePr != 48 & fires_ns$FireCause == 'Human' ~ "Human-caused wildfire")
n <- highlight_key(fires_ns, ~NEWCAT)
p1 <- 
  ggplot(n, aes(x=plot_date, fill = NEWCAT)) + 
  geom_density(alpha=0.7, color=NA) +
  scale_x_date(date_breaks = "months", date_labels = "%b") +
  scale_y_continuous(position = "right") +
  theme(plot.background = element_rect(fill='gray20'), 
        panel.background = element_rect(fill='gray20'), 
        axis.text.y = element_text(colour='gray98', size=16),
        axis.text.x = element_text(colour='gray98', size=10),
        panel.grid.major = element_blank(),   
        panel.grid.minor = element_blank(),
        panel.grid.major.y = element_line(size=1.2, color="gray"),
        axis.ticks.x = element_line(colour='gray98'),
        plot.margin=unit(c(0,2,0,0), 'cm')) +
    xlab("") +
     ylab("Density of fire occurences") +
    scale_fill_manual(values=c("#FF5733", "#ffff00","#56B4E9"))

  

p2<-ggplot(n, aes(y=FIRE_YEAR)) +
  geom_hline(yintercept = seq(1992, 2015, by = 1), color = "gray", size = 0.04) +
  scale_size_area(max_size = 15, guide = FALSE) +
  scale_x_date(date_breaks = "months", date_labels = "%b") +
  scale_y_reverse(limits = c(2015,1992), breaks = c(2010,2005,2000,1995)) +
  xlab("") +
  ylab("") +
  theme(plot.background = element_rect(fill='gray20'), 
        panel.background = element_rect(fill='gray20'), 
        axis.text.y = element_text(colour='gray98', size=16),
        axis.text.x = element_text(colour='gray98', size=10),
        panel.grid.major = element_blank(),   
        panel.grid.minor = element_blank(),
        panel.grid.major.y = element_line(size=1.2, color="gray"),
        axis.ticks.x = element_line(colour='gray98'),
        legend.text = element_text(colour='gray98'),
        legend.title = element_blank(),
        legend.background = element_rect(fill='gray20'),
        plot.margin=unit(c(0,2,0,0), 'cm')) +
  geom_point(aes(size = GISAcres*2, x = plot_date, color = NEWCAT, group=1, text=paste('Name:', IncidentNa, '<br>Type:', NEWCAT, '<br>Date:', format(StartDate, "%b %d %Y"), '<br>Duration:', duration, "days", "<br>Size:", format(round(as.numeric(GISAcres), 0), big.mark=","), "acres")), alpha=0.5) +
  scale_color_manual(values=c("#FF5733", "#ffff00","#56B4E9"))

p1g<-ggplotly(p1, tooltip=FALSE)
p2g<-ggplotly(p2, tooltip='text')

subplot(p2g, p1g, nrows=2) %>% hide_legend() %>% highlight("plotly_hover")
```




